Le titre

### 2022-09-15 Manuscrit Thèse/chapter 1/Chapter 1.tex
hgutidoe

For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one. For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach 50\% of the total NV concentration \citep{grezes2015storage}

@@

Fig. \ref{NV electrons}-b) shows the predicted molecular orbitals of the NV centers, initially computed as linear combination of the $sp^3$ orbitals respecting the $C_{3v}$ symmetry of the defect \citep{loubser1978electron}. This model has since been confirmed by ab initio computation and experimental observations \citep{doherty2013nitrogen}. In particular ab initio computations have shown that the molecular orbitals of the NV center are mostly localized on the nearest neighbors of the vacancy \citep{gali2008ab}. Ab initio computation has also determined that the $a'_1$ lies in the valence band and is therefore often ignored.

@@

Thanks to the discrete electronic transitions inside the diamond bad gap, the NV center is fluorescent: it can absorb light from $450-630$ nm and re-emits light in the $630-800$ nm band.

@@

\medskip
Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages: thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and close to 1 quantum efficiency \citep{kurtsiefer2000stable}, they are very bright. With proper nanophotonics engineering (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}. The NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}. There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}. This limits the potential of NV centers for quantum communication protocols that rely on a deterministic photon source \citep{bradac2019quantum}.

@@

Because $\gamma_n \ll \gamma_e$, the hyper-fine structure of the NV center can simply be thought of as a constant energy offset. This explains why for most intent and purposes the hyper-fine structure is ignored, along the electric and strain dependencies, which leads to the simpler spin Hamiltonian:

@@

\medskip
The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}. Part of the reason is the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}. However the main advantage of the NV center spin compared to other spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.

@@

Fig. \ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal. We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian. For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).

@@

\section{Measuring the NV center spin properties}

One of the most important characterization of qubits is their relaxation times. Indeed, the potential applications of most qubits in quantum information or quantum sensing protocols is ultimately limited by their relaxation rates \citep{de2021materials, degen2017quantum}.

The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ corresponding to the relaxation time of the spin populations to their thermal equilibrium, and the transverse relaxation time $T_2$ (following the NMR nomenclature) corresponding to the relaxation of the spin state coherences (non diagonal elements of the density matrix).

The transverse relaxation time, also referred to as coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$. The $T_2^*$ time includes the inhomogeneous dephasing processes such as the in homogeneity in space (when multiple qubits are probed) and time of the external fields. The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}. While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used. More careful nomenclatures will label the $T_2$ times $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials}. The relaxation times are ordered as follows:
\begin{equation}
T_2^* < T_2 < 2 T_1.
\end{equation}

The relaxation times of NV centers depend in large part on their environment, and can therefore vary by several order of magnitude depending on the sample used. The following paragraphs will cover the measurement processes of the relaxation times and the typical values for the samples used in this manuscript.

\subsection{$T_1$ measurement}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/T1_basique}
\caption{Basic $T_1$ measurement scheme. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink for low laser intensity. The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
\label{T1 basique}
\end{figure}

@@

The sequence shown in Fig.\ref{T1 basique} can be improved by implanting common mode rejection technique described in Fig.

@@

\subsection{$T_2^*$ measurement: Ramsey interferometry}
Parler aussi d'ODMR et pourquoi pas un exemple de différence entre 2 samples clean et pas clean

>>>>

For samples in high excess of [N$_s$] such as type Ib HPHT diamonds, the question of the charge state of the NV center is of relatively low importance as the NV$^-$ state is the predominant one. For sample with a lower [N$_s$] content and a higher [NV]/[N$_s$] ratio however, one must be more careful with the optical intensity and wavelength used as the concentration of NV$^0$ can reach up to 50\% of the total NV concentration \citep{grezes2015storage}

@@

Fig. \ref{NV electrons}-b) shows the predicted molecular orbitals of the NV centers, initially computed as linear combination of the $sp^3$ orbitals respecting the $C_{3v}$ symmetry of the defect \citep{loubser1978electron}. This model has since been confirmed by ab initio computation and experimental observations \citep{doherty2013nitrogen}. In particular ab initio computations have shown that the molecular orbitals of the NV center are mostly localized on the nearest neighbors of the vacancy \citep{gali2008ab}. Ab initio computation has also determined that the $a'_1$ level lies in the valence band and is thus often ignored.

@@

Because the NV center has discrete electronic transitions inside the diamond bad gap, it fluoresces: it can absorb light from $450-650$ nm and re-emits light in the $600-800$ nm band.

@@

\medskip
Compared to similar solid state photon sources (other color centers, quantum dots, dye molecules...), the NV center offer several advantages: thanks to the $\sim 11\ \rm ns$ lifetime of the excited state and close to 1 quantum efficiency \citep{kurtsiefer2000stable}, they are very bright. With proper nanophotonics engineering (solid immersion lenses or diamond waveguide) the PL collected from a single NV center can reach a few $10^6$ photons/s \citep{schroder2016quantum}. The NV center is also extremely photostable under green excitation, and does not show any photobleaching \citep{brouri2000photon}. There are however drawbacks in using NV centers for photonics applications, the main one being that $\leq 4\%$ of the photons emitted by the NV centers are in the ZPL (low Debye Waller factor), even at cryogenic temperature \citep{johnson2015tunable}. This limits the potential of NV centers for quantum communication protocols that rely on a deterministic photon source \citep{bradac2019quantum}.

@@

For most intent and purposes, the hyper-fine structure of the NV center can simply be thought of as an energy offset, and is thus often ignored. The electric and strain dependencies of the electronic Hamiltonian are also often ignored as they are of second order compared to the magnetic field dependency. This leads to the simpler spin Hamiltonian:

@@

\bigskip
The NV center's spin usage as a qubit is prevalent in quantum communication \citep{wehner2018quantum}, quantum computing \citep{de2021materials} and quantum sensing \citep{degen2017quantum}. Part of the reason are the overall long relaxation times (lifetime $T_1$ and coherence time $T_2$) of the order of a few ms at room temperature \citep{balasubramanian2009ultralong}, as well as the ability to address individual nuclear spins with even longer relaxation times \citep{awschalom2018quantum}. However the main advantage of the NV center spin compared to other solid state spin qubits is its ability to be polarized and read-out optically thanks to the properties detailed in the following section.

@@

Fig. \ref{ODMR 8 classes}-b) represent the 4 possible orientation of the N$-$V axis in a diamond single crystal. We refer to these orientations as ``classes" since NV centers belonging to the same class will perceive the same transverse and longitudinal magnetic field, and thus will have the same spin Hamiltonian. For large NV ensembles (a confocal spot can contain more than $10^9$ NV centers in dense ensemble), we can consider that each class is equally represented \footnote{It is possible to create NV centers with a preferential orientation \citep{lesik2014perfect} but this requires specific conditions during the growth process of the diamond} and that each class is responsible for two lines in the ODMR spectrum (one for the $\ket{0} \to \ket{-1}$ transition and one for the $\ket{0} \to \ket{+1}$ transition).

@@

\section{Relaxation times of the NV center spin}

This section focuses on the dynamics of the NV spin, and more specifically on the various relaxation times. Relaxation times are very important properties of spin qubits as they often are the limiting factor for quantum information or quantum sensing potential applications \citep{de2021materials, degen2017quantum}.

The relaxation times are generally decomposed between the longitudinal relaxation time $T_1$ which corresponds to the characteristic time of the return to thermal equilibrium for the populations (diagonal elements of the density matrix), and the transverse relaxation time $T_2$\footnote{We follow here the NMR nomenclature} which corresponds to the relaxation of the spin state coherences (non diagonal elements of the density matrix).

The transverse relaxation time, also called coherence time or dephasing time, is often further decomposed between the inhomogeneous dephasing time $T_2^*$ and the homogeneous dephasing time $T_2$. The $T_2^*$ time includes inhomogeneous dephasing processes such as the inhomogeneity in space (when multiple qubits are probed) and time of the external fields. The $T_2$ time corresponds to the coherence time where some of these inhomogeneities have been corrected, for instance with a Hanh echo sequence \citep{hahn1950spin} or a dynamical decoupling scheme \citep{naydenov2011dynamical}. While $T_2^*$ is an intrinsic property of the sample studied, $T_2$ depends on the rephasing scheme used. More careful nomenclatures will label the $T_2$ times $T_{2,\rm echo}$ or $T_{2,\rm DD}$ \citep{de2021materials} for instance. The relaxation times are ordered as follows:
\begin{equation}
T_2^* < T_2 < 2 T_1.
\end{equation}

The relaxation times of NV centers depend mainly on their environment, and can vary by several order of magnitude depending on the sample used. The following paragraphs will cover the measurement processes of the relaxation times and the typical values for the samples used in this manuscript.

\subsection{NV spin $T_1$}

\subsubsection{Measurement protocol}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/T1_basique}
\caption{Basic $T_1$ measurement scheme. (a) Pulse seqeunce used. (b) Experimental data obtained on sample CVD-pink for low laser intensity. The orange line corresponds to an exponential fit with time constant $\tau=4.0\ \rm ms$.}
\label{T1 basique}
\end{figure}

@@

The sequence shown in Fig.\ref{T1 basique} can be improved by implementing a common mode rejection technique described in Fig. \ref{T1 soustraction}-a). This protocol works as follow: first, a basic $T_1$ sequence is recorded, which we will denote $S_1$. Then the sequence is repeated with an additional $\pi$-pulse on one the NV transitions right before the readout. This $\pi$-pulse will project the remaining $\ket{0}$ population difference into a $\ket{\pm 1}$ state. The second readout is labeled $S_2$, and the final spin state readout is $S_1 -S_2$ which reject every contribution to the PL outside of the spin relaxation.

@@

Fig. \ref{T1 soustraction}-b) shows the experimental result of the $S_1$ and $S_2$ sequences on the same sample as for Fig. \ref{T1 basique}. The difference between the $S_1$ data in Fig. \ref{T1 soustraction}-b) and the the data in Fig. \ref{T1 basique}-b) comes from the laser intensity used. In Fig. \ref{T1 soustraction}, the laser intensity was high enough ($> 1\ \rm{mW}$) that photoionization between NV$^-$ and NV$^0$ becomes significant. Furthermore, for dense ensemble, a charge recombination mechanism operates even in the dark which is most likely due to charge tunneling between impurities \citep{choi2017depolarization, giri2018coupled, giri2019selective}. On the other hand, the laser intensity used in Fig. \ref{T1 basique} was small enough ($< 10\ \mu \rm{W}$) that the charge state state photodynamics contribution was negligible compared to the spin depolarization.

Fig. \ref{T1 soustraction}-c) Shows the subtraction of the $S_1$ and $S_2$ signal. Despite the complex shape of the initial $S_1$ and $S_2$ signals, $S_1-S_2$ is nicely fitted with an exponential decay similar to the results found in Fig. \ref{T1 basique}.

The photodynamics of the NV charge state is a complex subject which is still actively researched \citep{craik2020microwave, gorrini2021long}. The common mode rejection scheme described  here is often employed with dense ensemble of NV centers \citep{jarmola2012temperature, mrozek2015longitudinal, choi2017depolarization} as it allows to completely bypass the charge dynamics.

\subsubsection{Physical origin and typical values}

We distinguish three different origin for the the $T_1$ relaxation process:
\begin{itemize}
\item \textbf{Spin-phonon interaction:} the spin states are weakly coupled to the crystal phonon bath through spin-spin and spin-orbit interactions \citep{norambuena2018spin}, which results in a phonon-induced spin relaxation. The dominant relaxation mechanism depends on the  temperature regime, with the two-photon Raman process believed to be dominant at room temperature \citep{takahashi2008quenching, jarmola2012temperature} while the dominant process as at cryogenic temperature ($4\sim100$ K) is believed to be a two phonons Orbach process \citep{redman1991spin, norambuena2018spin}. The spin-phonon relaxation time strongly depends on the crystal lattice temperature, with relaxtion times of the order $T_1 \sim 5\ \rm ms$ at room temperature, and up to 8 hours at for mili-K temperatures \citep{astner2018solid}.

\item \textbf{Electric and magnetic field noises:} both electric and magnetic fields can directly excite the $\ket{0} \to \ket{\pm 1}$ spin transition \citep{udvarhelyi2018spin}. As a result, magnetic and electric field noise can reduce the spin $T_1$ if their spectrum overlap with the NV transitions. This is particularly important for near surface spins because of the strong noise caused by electronic defects on the surface \citep{sangtawesin2019origins}. The $T_1$ of near surface NV center can be as low as $\sim 10\ \mu$s, but the surface effect are only effective at very short range (within $\sim 10\ \rm nm$ of the surface).

\item \textbf{Resonant dipolar interaction with other paramagnetic impurities:} Dipolar interaction betwen two spins will be detailed more thoroughly in the following section. Dipole-dipole interaction between NV centers and surrounding spins is responsible for both transverse and longitudinal spin relaxation, with the added condition of energy matching (resonance) for longitudinal relaxation. Relaxation induced by NV-NV dipole-dipole coupling will be the object of chapter 3 and 4 while NV-other spin coupling will be the object of chapter 2. The $T_1$ time associated with dipole-dipole interaction strongly depends on the spin concentration, but it can be as low as $\sim 25 \mu$s \citep{hall2016detection}.
\end{itemize}

All the measurements shown in this manuscript were performed at room temperature. The spin-phonon interaction thus creates a baseline value for the NV $T_1$ time at $\sim 5\ \rm ms$. Every sample studied were at least a few $\mu$m thick and we will therefore neglect the surface noise contribution in our analyses. The dipole-dipole induced relaxation however is of crucial importance for the next chapters, where it will be further detailed. $T_1$ times between $0.1 \sim 1\ \rm ms$ were routinely observed for samples with high NV concentration.

\subsection{$T_2^*$ measurement: Ramsey interferometry}
Allez have fun
Parler aussi d'ODMR et pourquoi pas un exemple de différence entre 2 samples clean et pas clean

### 2022-09-15 Manuscrit Thèse/chapter 4/Chapter 4.tex
hgutidoe

To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes. This protocol however has an extremely low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic}.

>>>>

To our knowledge, the only known orientation-free protocol for DC magnetometry is to use the change in PL caused by the transverse field, which is a property that depends only weakly on the field orientation, especially with an ensemble of NV centers with multiple axes. This protocol however has an very low sensitivity and only had marginal use so far\citep{rondin2012nanoscale, tetienne2012magnetic, maletinsky2012robust, chapman2013background, jones2020selective}.
